---
title: "Portfolio Backtest Dashboard"
format:
  dashboard:
    theme: [default, theme.scss]
server: shiny
---

```{r}
#| context: setup
#| include: false
library(shiny)
library(quantmod)
library(PerformanceAnalytics)
source("portfolio_core.R")
```

## {.sidebar}

```{r}
textInput("symbols", "Symbols (comma-separated)", value = "VOO, TQQQ, SCHD")
textInput("weights", "Weights (comma-separated)", value = "40, 30, 30")
textInput("benchmark", "Benchmark", value = "SPY")
numericInput("years", "Years of history", value = 5, min = 1, max = MAX_YEARS, step = 1)
selectInput("rebal", "Rebalance", choices = REBAL_CHOICES, selected = "months")
actionButton("run", "Run Backtest", class = "btn-primary")
```

Dividends are reinvested (adjusted prices). If a fund's history is
shorter than the requested window, the analysis is trimmed to the
common date range.

Weights are relative and normalized automatically — e.g. weights
`1, 2` allocate 33% and 67%.

Up to 25 symbols and 30 years of history. Each symbol is one
download, so the cap is what keeps the wait bounded.

```{r}
#| context: server

parse_inputs <- eventReactive(input$run, {
  syms  <- parse_symbols(input$symbols)
  wts   <- parse_weights(input$weights)
  bench <- toupper(trimws(input$benchmark))
  errs  <- portfolio_input_errors(syms, wts, bench, input$years)
  validate(need(length(errs) == 0, paste(errs, collapse = "\n")))
  list(syms = syms, wts = normalize_weights(wts), bench = bench,
       years = input$years,
       rebal = input$rebal)
}, ignoreNULL = FALSE)  # runs once on startup with defaults

backtest <- reactive({
  p <- parse_inputs()
  from <- Sys.Date() - round(365.25 * p$years)

  get_prices <- function(t) {
    tryCatch(
      Ad(getSymbols(t, from = from, auto.assign = FALSE, warnings = FALSE)),
      error = function(e) NULL
    )
  }

  all_syms <- unique(c(p$syms, p$bench))

  # The downloads are sequential and block the session, so the wait is stated
  # instead of read as a frozen dashboard. The detail names the symbol being
  # waited on before the request goes out; the bar advances only once that
  # symbol is actually in hand, so it measures work done rather than work
  # started — a slow ticker shows as a stalled bar with its own name on it.
  raw <- withProgress(message = "Downloading prices", value = 0, {
    setNames(lapply(seq_along(all_syms), function(i) {
      incProgress(0, detail = all_syms[i])
      px <- get_prices(all_syms[i])
      incProgress(1 / length(all_syms))
      px
    }), all_syms)
  })

  failed <- names(raw)[vapply(raw, is.null, logical(1))]
  validate(need(length(failed) == 0,
    paste("Could not download:", paste(failed, collapse = ", "))))

  # unname() before do.call: a named list turns each ticker into a named
  # argument to merge.xts, and a symbol matching one of its parameters would be
  # eaten as that parameter rather than merged. Nothing reaches here in lower
  # case today — parse_symbols and the benchmark both toupper, and R's partial
  # matching is case-sensitive, so "ALL" is safe where "all" silently returns a
  # one-column result. That is a guarantee two edits away from not holding.
  prices <- do.call(merge, unname(raw))
  colnames(prices) <- all_syms
  prices <- na.omit(prices)
  validate(need(nrow(prices) > 30, "Not enough overlapping history for these symbols."))

  # Selection and colour are settled once, here, so every card below is drawing
  # from the same plan rather than each re-deriving its own.
  c(list(p = p, plan = display_plan(p$bench, p$syms, p$wts)),
    compute_backtest(prices, p$syms, p$wts, p$bench, p$rebal))
})

vb_note <- function(...) tags$span(class = "vb-note", ...)

chart_caption <- function(txt) tags$p(class = "chart-caption", txt)

output$vb_growth <- renderText({
  bt <- backtest()
  fmt_dollar(grew_to(bt$port))
})

output$vb_bench_growth <- renderText({
  bt <- backtest()
  sprintf("%s (%s)", fmt_dollar(grew_to(bt$bench)), bt$p$bench)
})

output$vb_cagr <- renderUI({
  bt <- backtest()
  cagr  <- ann_return(bt$port, bt$scale)
  bench <- ann_return(bt$bench, bt$scale)
  d <- delta_vs_benchmark(cagr, bench)
  tagList(fmt_pct(cagr),
    vb_note("Avg. yearly growth — ",
            tags$span(class = d$class, d$text),
            sprintf(" vs %s at %s", bt$p$bench, fmt_pct(bench))))
})

output$vb_sharpe <- renderUI({
  bt <- backtest()
  s <- ann_sharpe(bt$port, bt$scale)
  tagList(fmt_sharpe(s), vb_note(sharpe_caption(s)))
})

output$vb_maxdd <- renderUI({
  bt <- backtest()
  dd <- ann_maxdd(bt$port)
  tagList(fmt_pct(-dd),
    vb_note(sprintf("Biggest peak-to-trough loss — %s at the peak fell to %s",
                    fmt_dollar(START_CAPITAL),
                    fmt_dollar(START_CAPITAL * (1 - dd)))))
})

output$vb_period <- renderText({
  bt <- backtest()
  sprintf("%s – %s", fmt_month_year(start(bt$port)), fmt_month_year(end(bt$port)))
})

output$perf_caption <- renderUI({
  bt <- backtest()
  chart_caption(sprintf("%s · rebalanced %s · benchmark %s",
    paste0(round(bt$p$wts * 100, 1), "% ", bt$p$syms, collapse = " / "),
    rebal_label(bt$p$rebal), bt$p$bench))
})

# The card header already names each chart, so the plots draw no title of their
# own — the caption below carries the parameters instead, which is content
# rather than a second copy of the header.
output$perf_chart <- renderPlot({
  bt <- backtest()
  cols <- bt$plan$colors
  charts.PerformanceSummary(bt$combined,
    wealth.index = TRUE, main = "", legend.loc = "topleft",
    colorset = unname(cols[colnames(bt$combined)]),
    lwd = 2, element.color = GRID_COLOR,
    cex.axis = 0.85, cex.legend = 0.9, cex.lab = 0.85,
    main.timespan = FALSE)   # the Period value box already states the window
}, bg = "transparent", alt = function() {
  bt <- backtest()
  perf_chart_alt(bt$port, bt$bench, bt$p$bench)
})

output$fund_caption <- renderUI({
  bt <- backtest()
  shown <- bt$plan$shown
  chart_caption(if (length(shown) < length(bt$p$syms))
    sprintf("$1 in each holding on its own. Showing the %d largest of %d positions — the palette has %d distinct colours.",
            length(shown), length(bt$p$syms), length(SERIES_SLOTS))
  else "$1 in each holding on its own, dividends reinvested.")
})

output$fund_chart <- renderPlot({
  bt <- backtest()
  shown <- bt$plan$shown
  cols <- bt$plan$colors
  chart.CumReturns(bt$returns[, shown, drop = FALSE],
    wealth.index = TRUE, legend.loc = "topleft", main = "",
    colorset = unname(cols[shown]),
    lwd = 2, element.color = GRID_COLOR,
    cex.axis = 0.85, cex.legend = 0.9, cex.lab = 0.85,
    main.timespan = FALSE)
}, bg = "transparent", alt = function() {
  bt <- backtest()
  fund_chart_alt(bt$returns, bt$plan$shown)
})

# A colour chip per row ties each table row to its line in the charts, so the
# legend and the table are the same key. The name stays in text ink — the chip
# beside it carries the identity.
output$stats_table <- renderTable({
  bt <- backtest()
  tbl <- build_stats_table(bt$port, bt$returns, bt$p$bench, bt$p$syms, bt$scale)
  cols <- bt$plan$colors
  out <- cbind(" " = entity_chip(rownames(tbl), cols[rownames(tbl)]), tbl)
  names(out)[1] <- CHIP_COLUMN_HEADER
  out
}, rownames = FALSE, sanitize.text.function = identity)

# The calendar is polarity data — gains against losses — so it gets a diverging
# tint (blue ↔ red, neutral at zero) instead of a uniform wall of numerals. The
# sign is always printed, so the colour reinforces and never carries alone.
output$calendar_table <- renderUI({
  bt <- backtest()
  cal <- table.CalendarReturns(bt$port)
  last <- ncol(cal)                       # annual total, named after the series
  month_cap <- tint_cap(unlist(cal[, seq_len(last - 1)]))
  year_cap  <- tint_cap(cal[[last]])

  cell <- function(v, cap, total) {
    if (is.na(v)) return(tags$td(class = "cal-cell", ""))
    tags$td(class = if (total) "cal-cell cal-total" else "cal-cell",
            style = paste0("background:", return_tint(v, cap)),
            sprintf("%.1f", v))
  }

  # Month headers come from MONTH_ABB, not from table.CalendarReturns' own
  # colnames — those are built with %b and arrive localized ("Mrz", "Okt").
  tags$table(class = "table calendar",
    tags$thead(tags$tr(tags$th(""),
      lapply(c(MONTH_ABB[seq_len(last - 1)], "Year"),
             function(m) tags$th(scope = "col", m)))),
    tags$tbody(lapply(seq_len(nrow(cal)), function(i)
      tags$tr(tags$th(scope = "row", rownames(cal)[i]),
        lapply(seq_len(last), function(j)
          cell(cal[i, j], if (j == last) year_cap else month_cap, j == last))))))
})

output$dd_table <- renderTable({
  bt <- backtest()
  build_drawdown_table(bt$port, bt$scale)
})
```

## Column {width=20%}

```{r}
#| content: valuebox
#| title: "$10,000 grew to"
#| icon: cash-stack
#| color: primary
textOutput("vb_growth")
```

```{r}
#| content: valuebox
#| title: "In benchmark it grew to"
#| icon: bank
#| color: light
textOutput("vb_bench_growth")
```

```{r}
#| content: valuebox
#| title: "CAGR"
#| icon: graph-up-arrow
#| color: light
uiOutput("vb_cagr")
```

```{r}
#| content: valuebox
#| title: "Sharpe"
#| icon: speedometer2
#| color: light
uiOutput("vb_sharpe")
```

```{r}
#| content: valuebox
#| title: "Max Drawdown"
#| icon: graph-down-arrow
#| color: light
uiOutput("vb_maxdd")
```

```{r}
#| content: valuebox
#| title: "Period"
#| icon: calendar3
#| color: light
textOutput("vb_period")
```

## Column {width=40%}

### Row

```{r}
#| title: "Performance vs Benchmark"
plotOutput("perf_chart")
uiOutput("perf_caption")
```

### Row

```{r}
#| title: "Per-Holding Growth"
plotOutput("fund_chart")
uiOutput("fund_caption")
```

## Column {width=40%}

### Row

```{r}
#| title: "Key Statistics"
tableOutput("stats_table")
```

### Row

```{r}
#| title: "Monthly Returns (%)"
uiOutput("calendar_table")
```

### Row

```{r}
#| title: "Worst Drawdowns"
tableOutput("dd_table")
```
